---
title: "Premier League Fantasy Football"
author: "Barry Bullas MBE & Poppy Bullas"
date: "30/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

list_of_packages <- c("rvest", "dplyr", "httr", "xml2", "jsonlite", "DT")
new_packages <- list_of_packages[!(list_of_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)

#Load required packages
lapply(list_of_packages, library, character.only = TRUE)
```

# Premier League Fantasy Football 2019/20
The [Premier League Fantasy Football website](https://fantasy.premierleague.com/statistics) has a useful API feature that can be used to return data on players and teams.

This project is my attempt to retrieve this data, in the hope that i will better be able to predict which players i should transfer onto my team. I used [this post by Paddy KG](https://medium.com/@YourMumSaysWhat/how-to-get-data-from-the-fantasy-premier-league-api-4477d6a334c3) as a starting point.  The urls in his post have changed slightly this year but it helped me figure them out
 
## Player stats
The player stats include the ICT index, developed specifically to give a verdict on a player as an FPL asset.  You can read [How the ICT Index in Fantasy works](https://www.premierleague.com/news/65567).

This page [https://fantasy.premierleague.com/api/bootstrap-static/](https://fantasy.premierleague.com/api/bootstrap-static/) contains the stats for all players. 

Each player is given a unique id in the dataset.  If you want to identify the id of a specific player then you can search [https://fantasy.premierleague.com/api/bootstrap-static/](https://fantasy.premierleague.com/api/bootstrap-static/) in your browser for the name.  For example searching if you search for **Salah** in the data you would see that his player id is **191**.

    "id":191,"in_dreamteam":true,"news":"","news_added":null,"now_cost":125,"photo":"118748.jpg","points_per_game":"10.0","second_name":"Salah",

With the player id you can go to [https://fantasy.premierleague.com/api/element-summary/191/](https://fantasy.premierleague.com/api/element-summary/191/) to retrieve the players stats.  By simply replacing 191 with the id of any other player, you can view that players stats.

```{r}
static_url <-
  'https://fantasy.premierleague.com/api/bootstrap-static/'

#This text is passed to the GET using user_agent.  It allows those running the API to contact you should your requests be causing issues
user_text <-
  'my@email.address this is a personal project to test retrieving data from an api'

#Get the result
get_result_static <- GET(static_url, user_agent(user_text))

# Check http_type() of resp_json
http_type(get_result_static)

# Parse returned text with fromJSON()
all_stats_json <- fromJSON(content(get_result_static, as = "text"), simplifyDataFrame = TRUE)

# Player stats
players_df <- all_stats_json[["elements"]]

# Player types
player_type_df <- all_stats_json[["element_types"]]
player_type_df <- player_type_df[ ,c ("id", "plural_name_short")] %>%
  rename("player_type_id" = "id", "player_type" = "plural_name_short" )

# Team stats
team_df <- all_stats_json[["teams"]]

# Subset the id and team name to join to player stats
team_df <- team_df[,c("id", "name")] %>%
  rename ("team_name" = "name")

# Event stats
events_df <- all_stats_json[["events"]]

# Add Team Names to player stats
players_df <- merge(x = players_df, y = team_df, by.x = "team", by.y="id", all.x=TRUE)

# Add player types
players_df <- merge(x = players_df, y = player_type_df, by.x = "element_type", by.y="player_type_id", all.x=TRUE)
players_df <- players_df[order(players_df$id),]

#field selection
fields <- c("id", 
            "first_name",
            "second_name",
            "player_type",
            "team_name",
            "now_cost",
            "total_points",
            "points_per_game",
            "status",
            "bps",
            "influence",
            "creativity",
            "threat",
            "ict_index",
            "news",
            "news_added",
            "photo",
            "bonus",
            "value_form",
            "value_season",
            "goals_scored",
            "assists",
            "clean_sheets",
            "own_goals",
            "penalties_saved",
            "penalties_missed",
            "yellow_cards",
            "red_cards",
            "saves",
            "bonus",
            "element_type",
            "team")

# Remove columns not needed
players_df <- players_df[ ,fields]

# Add on url for images - **Not currently working - access denied**
photo_url <- 'https://platform-static-files.s3.amazonaws.com/premierleague/photos/players/110x140/'
players_df["photo"] <- paste0(photo_url, players_df[["photo"]])

# Update cost
players_df["now_cost"] <- players_df[["now_cost"]] * 100000


# Save it to csv
write.csv(players_df, 'data/player_stats.csv')

datatable(players_df, rownames= FALSE, options = list(autoWidth = FALSE, pageLength=5))



```


# Loop through all players

```{r }
# 
# 
# player_base_url <-
#   'https://fantasy.premierleague.com/api/element-summary/'
# 
# for (i in 1:10) {
#   #Create the URL for the next id
#   url <- paste0(player_base_url, i, '/')
#   
#   # Send a GET request to url
#   get_player_result <- GET(url, user_agent(user_text))
#   
#   # Delay for 1 seconds between requests
#   # This is helpful when making a lot of calls to an api
#   # IT ensures you don't overload the server causing your requests to be rejected
#   Sys.sleep(1)
#   
#   
#   #Check the result returned sucessfully
#   if (http_error(get_player_result)) {
#     warning("The request for player id "+i + " failed")
#   } else {
#     #Do something with the data
#     print(paste('GET sucessfully returned for player id ',  i))
#   }
# }
```

